---
title: "MERR - Bio Project"
author: "Adib HABBOU - Alae KHIDOUR"
date: "2022-11-07"
output: pdf_document
---

# Data Exploration

## Data Import

```{r}
load(file = "project_park.RData")
```

## Data Size

```{r}
dim(geno.df)
```

```{r}
dim(pheno.df)
```

```{r}
dim(Xmat)
```

## Phenotypes names

```{r}
names(pheno.df)
```

## Choice of target variables

```{r}
# Plant.height
i = 14
y = matrix(pheno.df[,i])
yname = names(pheno.df)[i]
summary(y)
```

```{r}
hist(y, main = "Frequency of plant height", xlab = "Height (cm)", col = "chartreuse3")
```

```{r}
# Flag.leaf.length
i = 10
y = matrix(pheno.df[,i])
yname = names(pheno.df)[i]
summary(y)
```

```{r}
hist(y, main = "Frequency of flag leaf length", xlab = "Length (cm)", col = "chartreuse3")
```

```{r}
# Flag.leaf.width
i = 11
y = matrix(pheno.df[,i])
yname = names(pheno.df)[i]
summary(y)
```

```{r}
hist(y, main = "Frequency of flag leaf width", xlab = "Width (cm)", col = "chartreuse3")
```

## Sample of data

```{r}
Xmat[1:5,1:5]
```

```{r}
geno.df[1:5,1:10]
```

# Data Preparation

## Plot of NA

```{r}
library(naniar)
library(ggplot2)
data<- pheno.df[,c(10,11,14,5,9,7,37,38,36,4,33,34,6,32,13,12,3,16,20,21,22,23)]
a <- c(rep("black", 3), "firebrick1","firebrick1","firebrick1", rep("black", length(data)-6))
gg_miss_var(data) +
  labs(y = "Number of missing values") + 
  ggtitle("Number of missing values for each variable") +
  theme(axis.text.y = element_text(colour = a))
```

## Imputation of NA in target variables

```{r}
Y <- pheno.df["Plant.height"]
mean <- mean(Y[!is.na(Y)])
Plant.height.is.missing <- vector(length = nrow(Y))
for (i in 1:nrow(Y))
{
  if (is.na(Y[i,1])) 
  {
    Plant.height.is.missing[i] <- 1
    Y[i,1] <- mean
  }
}
Plant.height <- cbind(Y,Plant.height.is.missing)
knitr::kable(head(Plant.height))
```

```{r}
Y <- pheno.df["Flag.leaf.length"]
mean <- mean(Y[!is.na(Y)])
Flag.leaf.length.is.missing <- vector(length = nrow(Y))
for (i in 1:nrow(Y))
{
  if (is.na(Y[i,1])) 
  {
    Flag.leaf.length.is.missing[i] <- 1
    Y[i,1] <- mean
  }
}
Flag.leaf.length <- cbind(Y,Flag.leaf.length.is.missing)
knitr::kable(head(Flag.leaf.length))
```

```{r}
Y <- pheno.df["Flag.leaf.width"]
mean <- mean(Y[!is.na(Y)])
Flag.leaf.width.is.missing <- vector(length = nrow(Y))
for (i in 1:nrow(Y))
{
  if (is.na(Y[i,1])) 
  {
    Flag.leaf.width.is.missing[i] <- 1
    Y[i,1] <- mean
  }
}
Flag.leaf.width <- cbind(Y,Flag.leaf.width.is.missing)
knitr::kable(head(Flag.leaf.width))
```

## Imputation of NA in Xmat

```{r}
X <- as.data.frame(Xmat)
dim(X)
```

```{r}
sum(is.na(X))
```

```{r}
X_no_NA <- t(na.omit(t(X)))
dim(X_no_NA)
```

```{r}
X_NA <- X[,-X_no_NA]
dim(X_NA)
```

```{r}
rows_to_drop <- c()
for (i in 1:nrow(X))
{
  if (mean(is.na(X[i,])) * 100 > 10) rows_to_drop <- c(rows_to_drop, i)
  if (mean(X[i,][!is.na(X[i,])] == rep(2, length(X[i,][!is.na(X[i,])]))) * 100 < 0.1) 
    rows_to_drop <- c(rows_to_drop, i)
}
```

```{r}
new_X <- X[-rows_to_drop,]
```

```{r}
target <- cbind(Plant.height, Flag.leaf.length, Flag.leaf.width)
dim(target)
```

```{r}
new_target <- target[-rows_to_drop,]
dim(new_target)
```

```{r}
write.csv(new_target, "bio_target.csv")
```

```{r}
library(tidyr)
for (j in 1:ncol(new_X))
{
  mean <- mean(new_X[,j], na.rm = TRUE)
  if (mean > 1) new_X[,j] <- replace_na(new_X[,j], 2)
  else new_X[,j] <- replace_na(new_X[,j], 0)
}
```

```{r}
write.csv(new_X, "bio_data.csv")
```

## Study of correlation

```{r}
library(corrplot)
library(dplyr)
```

```{r}
bio_data <- read.csv("bio_data.csv")
```

```{r}
corr_simple <- function(data, sig)
{
  df_cor <- data %>% mutate_if(is.character, as.factor)
  df_cor <- df_cor %>% mutate_if(is.factor, as.numeric)
  
  corr <- cor(df_cor)
  corr[lower.tri(corr,diag=TRUE)] <- NA 
  corr[corr == 1] <- NA
  
  corr <- as.data.frame(as.table(corr))
  corr <- na.omit(corr) 
  corr <- subset(corr, abs(Freq) > sig) 
  corr <- corr[order(-abs(corr$Freq)),]
  
  print(corr)
  
  mtx_corr <- reshape2::acast(corr, Var1~Var2, value.var="Freq")
  corrplot(mtx_corr, is.corr=FALSE, tl.col="black", na.label=" ", 
           tl.cex = 0.25, mar=c(0,0,1,0), title="Correlation Plot")
}
```

```{r}
corr_simple(bio_data[, c(1000:1300, 28800:29600, 35500:36000)], 0.998)
```

# Data Modeling

## Data Importation

```{r}
bio_target <- read.csv("bio_target.csv")
```

```{r}
Plant.height <- bio_target$Plant.height
plant_height <- cbind(bio_data, Plant.height)
```

## Simple Regression

```{r}
plant_height <- cbind(bio_data[,2:12300], Plant.height)
plant_height_simple <- lm(formula = Plant.height ~ ., data = plant_height)
na_var <- c()
index <- 0
for (x in plant_height_simple$coefficients)
{
  index <- index + 1
  if (is.na(x)) na_var <- c(na_var, index)
}
plant_height_1 <- bio_data[,2:12300][,-(na_var-1)]
```

```{r}
plant_height <- cbind(bio_data[,12300:24600], Plant.height)
plant_height_simple <- lm(formula = Plant.height ~ ., data = plant_height)
na_var <- c()
index <- 0
for (x in plant_height_simple$coefficients)
{
  index <- index + 1
  if (is.na(x)) na_var <- c(na_var, index)
}
plant_height_2 <- bio_data[,12300:24600][,-(na_var-1)]
```

```{r}
plant_height <- cbind(bio_data[,24600:36902], Plant.height)
plant_height_simple <- lm(formula = Plant.height ~ ., data = plant_height)
na_var <- c()
index <- 0
for (x in plant_height_simple$coefficients)
{
  index <- index + 1
  if (is.na(x)) na_var <- c(na_var, index)
}
plant_height_3 <- bio_data[,24600:36902][,-(na_var-1)]
```

```{r}
plant_height_drop <- cbind(plant_height_1, plant_height_2, plant_height_3, Plant.height)
```

## Forward Regression

```{r}
plant_height_forward <- step(lm(Plant.height ~ 1, data = plant_height_drop), direction = "forward",
                             list(upper = lm(Plant.height ~ ., data = plant_height_drop)))
```

## Stepwise Regression

```{r}
plant_height_stepwise <- step(lm(Plant.height ~ 1, data = plant_height_drop), direction = "both",
                             list(upper = lm(Plant.height ~ ., data = plant_height_drop)))
```

## Variable Selection

```{r}
variables <- row.names(as.data.frame(plant_height_stepwise$coefficients))
data_selected <- data.frame(bio_data[variables[2]])
for (i in 3:length(variables))
{
  data_selected <- cbind(data_selected, bio_data[variables[i]])
}
```

## Ridge Regression

```{r}
library(glmnet)
```

```{r}
ridge <- glmnet(x = data_selected, y = Plant.height, alpha = 0, family = "gaussian")
plot(ridge, xvar = "lambda", label = FALSE, lwd = 0.5)
```

```{r}
cv_ridge <- cv.glmnet(as.matrix(data_selected), Plant.height, 
                      family = "gaussian", alpha = 0, type.measure = "mse")
plot(cv_ridge)
```

```{r}
ridge_lambda_min <- cv_ridge$lambda.min
ridge_min <- glmnet(x = data_selected, y = Plant.height, 
                    alpha = 0, family = "gaussian", lambda = ridge_lambda_min)
prediction <- predict(ridge_min, as.matrix(data_selected), type = "response")
R2_ridge <- 1 - sum((Plant.height - prediction)^2)/sum((Plant.height - mean(Plant.height))^2)
R2_ridge
```

## Lasso Regression

```{r}
lasso <- glmnet(x = data_selected, y = Plant.height, alpha = 1, family = "gaussian")
plot(lasso, xvar = "lambda", label = FALSE, lwd = 0.5)
```

```{r}
cv_lasso <- cv.glmnet(as.matrix(data_selected), Plant.height, 
                      family = "gaussian", alpha = 1, type.measure = "mse")
plot(cv_lasso)
```

```{r}
lasso_lambda_min <- cv_lasso$lambda.min
lasso_min <- glmnet(x = data_selected, y = Plant.height, 
                    alpha = 1, family = "gaussian", lambda = lasso_lambda_min)
prediction <- predict(lasso_min, as.matrix(data_selected), type = "response")
R2_lasso <- 1 - sum((Plant.height - prediction)^2)/sum((Plant.height - mean(Plant.height))^2)
R2_lasso
```

## Elastic Net

```{r}
elastic_net <- glmnet(x = data_selected, y = Plant.height, alpha = 0.5, family = "gaussian")
plot(elastic_net, xvar = "lambda", label = FALSE, lwd = 0.5)
```

```{r}
cv_elastic_net <- cv.glmnet(as.matrix(data_selected), Plant.height, 
                      family = "gaussian", alpha = 0.5, type.measure = "mse")
plot(cv_elastic_net)
```

```{r}
elastic_net_lambda_min <- cv_elastic_net$lambda.min
elastic_net_min <- glmnet(x = data_selected, y = Plant.height, 
                    alpha = 1, family = "gaussian", lambda = elastic_net_lambda_min)
prediction <- predict(elastic_net_min, as.matrix(data_selected), type = "response")
R2_elastic_net <- 1 - sum((Plant.height - prediction)^2)/sum((Plant.height - mean(Plant.height))^2)
R2_elastic_net
```

## Group-Lasso

```{r}
library(gglasso)
```

```{r}
group_lasso <- gglasso(x = as.matrix(data_selected), y = Plant.height, loss = "ls")
plot(group_lasso , xvar = "lambda", label = FALSE, lwd = 0.5)
```

```{r}
cv_group_lasso <- cv.gglasso(x = as.matrix(data_selected), y = Plant.height, 
                             group = group_lasso$group)
plot(cv_group_lasso)
```

```{r}
group_lasso_lambda_min <- cv_group_lasso$lambda.min
group_lasso_min <- gglasso(x = as.matrix(data_selected), y = Plant.height, 
                           loss = "ls", lambda = elastic_net_lambda_min)
prediction <- predict(group_lasso_min, as.matrix(data_selected), type = "link")
R2_group_lasso <- 1 - sum((Plant.height - prediction)^2)/sum((Plant.height - mean(Plant.height))^2)
R2_group_lasso
```

## Support Vector Regression

```{r}
library(e1071)
```

```{r}
svr <- svm(Plant.height ~ ., data = cbind(data_selected, Plant.height))
plot(svr, data = data_selected)
```

```{r}
prediction <- predict(svr, data_selected)
R2_svr <- 1 - sum((Plant.height - prediction)^2)/sum((Plant.height - mean(Plant.height))^2)
R2_svr
```

## Most Important Genetoypes

```{r}
coef <- coef(cv_lasso)
id_genes <- c()
for (i in 2:length(coef))
{
  if (coef[i] != 0) id_genes <- c(id_genes, i)
}
names <- row.names(coef)
genes <- c()
for (i in id_genes)
{
  genes <- c(genes, names[i])
}
genes
```


